A Large-scale Study of Polymorphism in R

### 2019-03-21 paper/main.tex
m

%
%
%
%
\section{The \texttt{genthat} Package}
\label{sec:genthat}

As we mentioned, getting our hands on ``front-line'' non-library R code is tricky: 
R is not like many other languages, where large systems are built up and stored in a repository such as GitHub.  
Instead, R programmers are liable to leverage a number of packages and work them into their data analysis pipelines.

Thankfully, we are somewhat able to make up for this lack of code.
CRAN has a policy that published packages are to be accompanied by at least one vignette, and typically by tests and examples showing off how the code is intended to be used.  
Nominally unrelated, the {\tt genthat} project~\cite{issta18} traces this example code to generate new unit tests which package designers may not have thought of.
As it happens, we can leverage {\tt genthat} to extract dynamic type information from running package test code, which presumably indicates how package code is intended to be used.

\AT{Draft}

@@

An unfortunate consequence of R's most common usage pattern is the lack of "deployed" R code. 
R is most often used for exploratory data analysis, and very little of this analysis code is made publicly available on e.g. GitHub.
The {\tt genthat} angle is thus desirable: 
With the scarcity of non-library code, examples of package usage is about as representative as one can hope for of the sort of code people will write in their analysis pipelines.

We can do a little bit better than merely package usage examples, though.
Available through CRAN is a list of {\it reverse dependencies} of packages: 
for some package {\tt p}, a list of all packages which require {\tt p} in some way.  
Our tracer is able to leverage and trace this code as well. 
When running a package, we also trace how all inherited functions are used, thus painting an even clearer picture of function usage.  
This additional data is considered alongside similar data collected for all other packages, as well as the data obtained by tracing the package's examples.

% Some prose we might want to pull from.
%First, we leverage {\tt genthat}'s tracer to trace function executions.  The
%idea here is that we don't really have access to non-library code written in
%R, as general use patterns are (possibly?) to write small scripts which
%analyze some bit of data and possibly visualize results.  The goal is likely
%not to build big working systems, instead to explore data with by writing
%and rewriting small scripts, ad infinitum.  In looking at package tests,
%vignettes, and examples, we are painting a picture of how the package
%designers intended their packages to be used, which we believe is a close
%approximation of what R users would do.

\section{The Method}

%
%
%
%
\subsection{Granularity of Type Information}

%
%
%
%
\subsection{Data Collection Pipeline}

As we mentioned in Section~\ref{sec:genthat}, we modify the {\tt genthat} tracer to collect and output the type information for all function invocations.

The high-level process for tracing one package is as follows:

@@

  	\item next, runnable code is extracted and run from package tests, examples, and vignettes, and \AT{what about rev deps}.

@@

	\item finally, after all test code has finished executing, we are left with a directory full of all of the unique type signatures that were observed while running the package code.

\end{enumerate}

@@

	\item an argument is {\it polymorphic in type} if it has been called with at least two distinct ``types'' according to the \code{typeof} function and our extra analysis.

	\item similarly, an argument is {\it polymorphic in class} if it has been called with at least two distinct classes according to the \code{class} function.

@@

%
%
%
%
%
%
\section{Results}
\label{sec:results}

@@

\AT{Draft}
Having built an understanding of our data collection and analysis pipeline, we will turn our attention to the results of our analysis.
We approach the data with the purpose of informing the design of a static type system for R.
To that end, we propose to begin with a very basic type system, the {\it Fine-Grained Type System} (FGTS).

>>>>

\newpage
\section{Methods}

@@

%
The high-level process for tracing one package is as follows:

@@

\item next, runnable code is extracted and run from package tests, examples,
  and vignettes.

@@

\item finally, after all test code has finished executing, we are left with
  a directory full of all of the unique type signatures that were observed
  while running the package code.
\end{enumerate}

@@

\begin{itemize}
\item an argument is {\it polymorphic in type} if it has been called with at
  least two distinct ``types'' according to the \code{typeof} function and
  our extra analysis.
\item similarly, an argument is {\it polymorphic in class} if it has been
  called with at least two distinct classes according to the \code{class}
  function.

@@

%
%

%
\subsection{Granularity of Type Information}

%
\section{Results}\label{sec:results}

@@

Having built an understanding of our data collection and analysis
pipeline, we will turn our attention to the results of our analysis.  We
approach the data with the purpose of informing the design of a static type
system for R.  To that end, we propose to begin with a very basic type
system, the {\it Fine-Grained Type System} (FGTS).
